---
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, fig.width=7, fig.height=2.5, fig.retina=2)
```

# Who needs `httpstat`?

`httpstat` (whether in [Python](https://github.com/reorx/httpstat) or [Go](https://github.com/davecheney/httpstat)) is all the rage these days with those Sillycanny Valley folks. We can do _so_ much better than text with R and we can work directly with what `httr` verb functions return:

```{r}
library(httr)
library(curl)
library(purrr)
library(dplyr)
library(tidyr)
library(stringi)
library(ggplot2)
library(hrbrmisc) # devtools::install_github("hrbrmstr/hrbrmisc")

httr_stat <- function(res) {
  
  if ("redirect" %notin% names(res$times)) {
    res$times <- c(redirect=0.00, res$times)
  }
  
  full_times <- c(res$times[1:5], contenttransfer=unname(res$times[6]-res$times[5]), res$times[6])
  start_pos <- c(0, unname(cumsum(c(0, full_times[2:6]))))
  end_pos <- start_pos + full_times
  stat <- c("Redirect", "Name Lookup", "Connect", "Pre-Transfer", "Start Transfer",
            "Content Transfer", "Total")
  
  df <- data_frame(stat, start_pos, end_pos, duration=full_times)
  
  df$start_pos[1] <- -df$end_pos[1]
  df$end_pos[1] <- 0
  df <- head(df, -1)
  
  df$stat <- factor(df$stat, levels=rev(df$stat))
  df$start_pos <- 100 * df$start_pos
  df$end_pos <- 100 * df$end_pos
  df$duration <- 100 * df$duration
  
  rng <- if (df$start_pos[1]<0) 1:6 else 2:6
  
  lims <- c(df$start_pos[1]*1.1, df$end_pos[6]*1.33)
  if (df$start_pos>0) lims[1] <- -0.5
  
  ggplot(head(df[rng,])) +
    geom_vline(data=df[2:6, ], aes(xintercept=end_pos), linetype="dotted", size=0.25, alpha=0.5) +
    geom_segment(aes(x=start_pos, xend=end_pos, y=stat, yend=stat),
                 arrow=arrow(ends="last", type="closed", length=unit(3, "points"))) +
    geom_label(aes(x=end_pos, y=stat, label=sprintf("(%4.3f ms)", duration)),
               hjust=0, family="Hind-Light", size=3, nudge_x=1, label.size=0) +
    geom_label(data=df[6,], aes(x=end_pos, y="Connect", 
                                label=sprintf("Total: %4.3f ms", end_pos)),
               hjust=0, family="Hind-SemiBold", size=3, nudge_x=0.5, label.size=0) +
    scale_x_continuous(expand=c(0,0.5), limits=lims, breaks=0, labels="0") +
    labs(x=NULL, y=NULL, title=NULL, subtitle=sprintf("Timings for %s", res$url)) +
    theme_hrbrmstr(grid="X", subtitle_family = "Hind") +
    theme(plot.subtitle=element_text(margin=margin(b=6))) -> gg
  
  print(gg)
  
  if (inherits(res, "response")) {
    
    walk(res$all_headers, function(x) {
      
      ah <- x[[3]]
      hnam <- stri_trans_totitle(names(ah))
      hval <- flatten_chr(ah)
      ahl <- max(map_int(hnam, nchar))
      
      cat(sprintf("%s: %s %s\n", stri_pad_left("STATUS", ahl), x[[2]], x[[1]]))
      cat(sprintf("%s: %s\n", stri_pad_left(hnam, ahl), hval), sep="")
      cat("\n", sep="")
      
    })
    
  } else {
    
    cat(rawToChar(res$headers), sep="")
    
  }
  
}
```

Since `httr` is configured to follow redirects, we have that data in the response and can use that in the output:

```{r}
httr_stat(GET("http://rud.is/b"))
```

Of course, it works fine when there aren't redirects:

```{r}
httr_stat(GET("http://rud.is/b/"))
```

And, it also works on raw `curl` requests:

```{r}
httr_stat(curl_fetch_memory("http://rud.is/b"))
```

```{r}
httr_stat(curl_fetch_memory("http://rud.is/b/"))
```

It's interesting to see how some sites are configured (and how well they perform):

```{r}
httr_stat(GET("http://r-project.org/"))
```

```{r}
httr_stat(GET("http://rstudio.com/"))
```

```{r}
httr_stat(GET("http://github.com/"))
```

```{r}
httr_stat(GET("http://stackoverflow.com/"))
```